---
title: "SocioDemographic"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
rm(list=ls()) #clears variables in global environment :-)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "//vu1file.it.vanderbilt.edu/VIEE/Shared/Water Security/Water Quality")
getwd() #automatically has it out of the RCode directory
#setwd("/vu1file.it.vanderbilt.edu/VIEE/Shared/Water Security")
```

####Packages needed
```{r}
#Include packages at top so they can all load at the start
library(tidyverse) #includes ggplot2 (for plotting), dplyr (piping), tidyr, readr, tibble, purr
library(reshape2) #for melt, dcast fns
library(broom) #convert into tidy dfs
library(dplyr)
library(tidyr)
library(readxl)
library(stringr)

StateList <- read.csv('../Data/RawData/StateList.csv')
StateList$State <- as.character(StateList$State)
```

###Urban Population
Data represents percentage of population that's urban in each state for a given census period

```{r}
urban <- read_excel('../Data/RawData/pop-urban-pct-historical.xlsx', sheet = "States", skip = 5)
colnames(urban)

urban_subset <- urban[c("Area Name","2010", "2000")] %>% #only keep the most pertinent columns
  filter(complete.cases(.))  #get rid of trailing NAs

#only keep the 48 contiguous states
urban_subset_48 <- merge(urban_subset, StateList, by.x= 'Area Name', by.y = "State", all.y = TRUE) #looks good - all 48 populated

urban_final <- urban_subset_48 %>%
  rename(UrbanPercent_2010 = `2010`) %>% #renaming columns for clarity
  rename(UrbanPercent_2000 = `2000`) %>% #renaming columns for clarity
  rename(State = `Area Name`) %>%
  mutate(UrbanGrowthRate = UrbanPercent_2010 - UrbanPercent_2000) %>% #calculating growth rate in urban between two years
  select(State, UrbanPercent_2010, UrbanGrowthRate) %>% #keep select columns
  melt(id = "State") %>%
  mutate(Units = "Percentage") %>% #add metadata
  mutate(Units = ifelse(variable == "UrbanPercent_2010", 2010, "2000-2010")) %>% #2010 data for urban percentage and 2000-2010 for growth rate data
  mutate(Notes = "Source: ICIP Data") #Full link: https://www.icip.iastate.edu/tables/population/urban-pct-states 

#write.csv(urban_final,'../Data/FormattedData/UrbanPercent_Growth.csv')
rm(urban, urban_48,urban_final,urban_subset,urban_subset_48)
```


###Population
Best to use the population estimate column for 2010-2017 for consistency sake

```{r}
population<-read_csv("../Data/RawData/population.csv", skip = 1)

population_subset <- population %>%
  select(Geography, matches("Population Estimate"))

#contiguous 48
population_subset_48 <-merge(population, StateList, by.x = "Geography", by.y = "State", all.y = TRUE)

#calculate population growth rate and clean up df
population_final <- population_subset_48 %>%
  rename(State = Geography) %>%
  select(State, matches("Population Estimate")) %>%
  melt(id = "State") %>%
  filter(variable %in% c("Population Estimate (as of July 1) - 2010","Population Estimate (as of July 1) - 2016")) %>%
  mutate(GrowthRate = ((value-value[variable == "Population Estimate (as of July 1) - 2010"]) / value[variable == "Population Estimate (as of July 1) - 2010"]) * 100 / 7) %>% #formula is ((now-past)/past)/#of yrs * 100%, http://pages.uoregon.edu/rgp/PPPM613/class8a.htm 
  filter(variable %in% "Population Estimate (as of July 1) - 2016") %>%
  melt(id = "State") %>%
  filter(!value %in% "Population Estimate (as of July 1) - 2016") %>%
  mutate(variable = ifelse(variable == "GrowthRate","Annual Population Growth Rate","Population Size")) %>%
  mutate(Year = ifelse(variable == "Annual Population Growth Rate", "2010-2016", 2016)) %>%
  mutate(Units = ifelse(variable == "Annual Population Growth Rate", "Percentage", "Count")) %>%
  mutate(Notes = "Source: US Census Bureau")

#write.csv(population_final,'../Data/FormattedData/PopulationSize_Growth.csv')
rm(population, population_final, population_subset, population_subset_48)
```


##Education
Extract percentage of state population with high school, undergrad, and grad education between 2011-2015


```{r}
education <- read_csv("../Data/RawData/education.csv") 

education_subset <- education %>%
  select(`GEO.display-label`,HC01_VC95, HC01_VC96, HC01_VC92) #contains 2011-2015 years for high school, undergrad, and grad; if 2006-2010 are also needed, include HC02_VC95, HC02_VC96, HC02_VC92

colnames(education_subset) <-education_subset[1, ] # making the first row the new column names 
education_subset <- education_subset [-1, ] # deleting the duplicate 

#only keep 48 contiguous 48 states
education_subset_48 <- merge(education_subset, StateList, by.x = "Geography", by.y = "State", all.y = TRUE) #all 48 there and populated :-)

#create final df
education_final <- education_subset_48 %>%
  select(-State_Abbreviation) %>%
  rename(State = Geography) %>%
  melt(id="State") %>%
  mutate(SplitString = str_split(variable, "([ - ])")) %>%
  mutate(variable = sapply(SplitString, "[[" , 7)) %>% #extract key word
  mutate(variable = ifelse(variable == "high", "Education: High School or higher",
                           ifelse(variable == "bachelor's", "Education: Bachelor's or higher", "Education: Graduate or Professional degree"))) %>% #rename sensibly
  mutate(Year = sapply(SplitString, "[[" , 1)) %>%
  select(-SplitString) %>%
  mutate(Units = "Percentage") %>%
  mutate(Notes = "Source: U.S. Census Bureau, American Community Survey 5-Year Estimates")
  
#write.csv(education_final,'../Data/FormattedData/Education.csv')
rm(education,education_final,education_subset,education_subset_48)
```

###Race and Ethnicity 
Note: hispanic or latino is an ethnicity whereas white/black are races

```{r}
RaceEthnicity <- read_csv("../Data/RawData/race_age.csv") 

RaceEthnicity_subset <- RaceEthnicity %>%
  select(`GEO.display-label`,HC01_VC94, HC01_VC88, HC01_VC95) #contains 2011-2015 years; if 2006-2010 are also needed, include HC02_VC94, HC02_VC88, HC02_VC95

colnames(RaceEthnicity_subset) <-RaceEthnicity_subset[1, ] # making the first row the new column names 
RaceEthnicity_subset <- RaceEthnicity_subset [-1, ] # deleting the duplicate 

#only keep 48 contiguous 48 states
RaceEthnicity_subset_48 <- merge(RaceEthnicity_subset, StateList, by.x = "Geography", by.y = "State", all.y = TRUE) #all 48 there and populated :-)
  
#Create final df
RaceEthnicity_final <- RaceEthnicity_subset_48 %>%
  select(-State_Abbreviation) %>%
  rename(State = Geography) %>%
  melt(id="State") %>%
  mutate(SplitString = str_split(variable, "([ - ])")) %>%
  mutate(variable = sapply(SplitString, "[[" , 17)) %>% #extract key word
  mutate(variable = ifelse(variable == "White", "White (Non-Hispanic-Latino)",
                           ifelse(variable == "Black", "Black (Non-Hispanic-Latino)", "Total Hispanic-Latino (of any race)"))) %>% #rename sensibly
  mutate(Year = sapply(SplitString, "[[" , 1)) %>%
  select(-SplitString) %>%
  mutate(Units = "Percentage") %>%
  mutate(Notes = "Source: U.S. Census Bureau, American Community Survey 5-Year Estimates")
  
#write.csv(RaceEthnicity_final,'../Data/FormattedData/RaceEthnicity.csv')
rm(RaceEthnicity,RaceEthnicity_final,RaceEthnicity_subset,RaceEthnicity_subset_48)
```

###Age
Extract median age and % of people over 62+ and % of ppl between 20-34 years

```{r}
Age <- read_csv("../Data/RawData/race_age.csv")

Age_subset <- Age %>%
  select(`GEO.display-label`,HC01_VC23, HC01_VC12, HC01_VC13,HC01_VC28) #contains 2011-2015 years; if 2006-2010 are also needed, include HC02_VC23, HC02_VC12, HC02_VC13, HC02_VC28

colnames(Age_subset) <-Age_subset[1, ] # making the first row the new column names 
Age_subset <- Age_subset [-1, ] # deleting the duplicate 

#only keep 48 contiguous 48 states
Age_subset_48 <- merge(Age_subset, StateList, by.x = "Geography", by.y = "State", all.y = TRUE) #all 48 there and populated :-)
  
#Create final df
Age_final <- Age_subset_48 %>%
  select(-State_Abbreviation) %>%
  rename(State = Geography) %>%
  melt(id="State") %>%
  mutate(SplitString = str_split(variable, "([ - ])")) %>%
  mutate(variable = sapply(SplitString, "[[" , 7)) %>% #extract key word
  mutate(variable = ifelse(variable == "Median", "Median Age",
                           ifelse(variable == "20", "20 to 24", 
                                  ifelse(variable == "25", "25 to 34", "62 and over")))) %>% #rename sensibly
  select(-SplitString) %>%
  spread(variable, value) %>%
  mutate(`20 to 24` = as.numeric(`20 to 24`)) %>%
  mutate(`25 to 34` = as.numeric(`25 to 34`)) %>%
  mutate(`20 to 34` = `20 to 24` + `25 to 34`) %>% #combine these two groups together
  melt(id = "State") %>%
  filter(!variable %in% c("20 to 24", "25 to 34")) %>%
  mutate(Year = "2011-2015") %>%
  mutate(Units = ifelse(variable == "Median Age", "Years", "Percentage")) %>%
  mutate(Notes = "Source: U.S. Census Bureau, American Community Survey 5-Year Estimates")
  
#write.csv(Age_final,'../Data/FormattedData/Age.csv')
rm(Age_final,Age,Age_subset,Age_subset_48)
```


###Labor


```{r}
Labor <- read_csv("../Data/RawData/labor.csv")

Labor_subset <- Labor %>%
  select(`GEO.display-label`, HC01_VC06, HC01_VC12,HC01_VC162, HC01_VC50, HC01_VC51, HC01_VC52, HC01_VC55, HC01_VC85) #contains 2011-2015 years; if 2006-2010 are also needed, include: HC02_VC06, HC02_VC12, HC02_VC162, HC02_VC50, HC02_VC51, HC02_VC52, HC02_VC55, HC02_VC85  

colnames(Labor_subset) <-Labor_subset[1, ] # making the first row the new column names 
Labor_subset <- Labor_subset [-1, ] # deleting the duplicate 

#only keep 48 contiguous 48 states
Labor_subset_48 <- merge(Labor_subset, StateList, by.x = "Geography", by.y = "State", all.y = TRUE) #all 48 there and populated :-)

#Columns are crazy weird so renaming ahead of time
colnames(Labor_subset_48) <- c("State","Employed (16 years and over)","Unemployment Rate","Below Poverty Level","Employed in Ag_Fishing_Mining",
                               "Employed in Construction", "Employed in Manufacturing","Employed in Transportation","Median Household Income", "State_Abbreviation")

#Create final df
Labor_final <- Labor_subset_48 %>%
  select(-State_Abbreviation) %>%
  melt(id="State") %>%
  mutate(Year = "2011-2015") %>%
  mutate(Units = ifelse(variable == "Median Household Income", "Dollars (Inflation-Adjusted)", "Percentage")) %>%
  mutate(Notes = "Source: U.S. Census Bureau, American Community Survey 5-Year Estimates")

#write.csv(Labor_final,'../Data/FormattedData/Labor.csv')
rm(Labor, Labor_final, Labor_subset, Labor_subset_48)
```

