---
title: "Governance and Physical Resources"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls()) #clears variables in global environment :-)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "//vu1file.it.vanderbilt.edu/VIEE/Shared/Water Security/Water Quality")
getwd() #automatically has it out of the RCode directory
#setwd("//vu1file.it.vanderbilt.edu/VIEE/Shared/Water Security")
```
####Packages needed
```{r}
#Include packages at top so they can all load at the start
library(tidyverse) #includes ggplot2 (for plotting), dplyr (piping), tidyr, readr, tibble, purr
library(reshape2) #for melt, dcast fns
library(broom) #convert into tidy dfs
library(readxl)
library(stringr)

StateList <- read.csv('../Data/RawData/StateList.csv')
StateList$State <- as.character(StateList$State)
```
#Governance

###Interstate compacts

```{r}
Interstate <- read_excel('../Data/RawData/InterstateCompacts.xlsx', 
                             sheet = "Sheet1", skip = 1)
Interstate$`Resolution (Click for Details)` <- NULL
Interstate <- Interstate[complete.cases(Interstate),] #drop leading NAs

Interstate_Categorical <- Interstate %>%
  count(`Participating State(s)`, sort=TRUE) %>%
  arrange(`Participating State(s)`)
  
#rename cols for clarity
colnames(Interstate_Categorical) <- c("State","value")
Interstate_Categorical$variable<-"Number of Interstate Water Compacts"

#make sure all/only the 48 states are on the list
FullInterstate <- merge(Interstate_Categorical, StateList, by = "State", all.y = TRUE)

#replace NA's with 0's
FullInterstate$value[is.na(FullInterstate$value)] <- 0 
FullInterstate$variable<-"Number of Interstate Water Compacts"

#write.csv(FullInterstate,"../Data/FormattedData/InterstateCompacts.csv")
rm(FullInterstate,Interstate,Interstate_Categorical)
```

###Nutrient Trading Programs
Binary variable (1=state=level trading program exists; 0=no trading program)

```{r}
Nutrient <- read_excel('../Data/RawData/NutrientTrading.xlsx', 
                             sheet = "Sheet1")
Nutrient$`Source: http://www.envtn.org/water-quality-trading/state-programs` <- NULL
colnames(Nutrient)<-c("State","value")
Nutrient$variable<-"State-Level Nutrient Trading Program"
Nutrient$Notes <- "Binary: 1 = state-level trading program exists"

Nutrient <- Nutrient %>%
  select(State,value,variable,Notes)

#write.csv(Nutrient,"../Data/FormattedData/NutrientTradingPrgrms.csv")
rm(Nutrient)
```

###Climate Change Planning
Discussion of Climate change in state water plan (Werner Svedin 2017), discussion of CC in hazard mitigation plan (Babcock 2013), state water agency practices as reported to EPA (EPA 2017) are all in the raw data. For simplicity, will just use the Werner and Svedin's analysis in the formatted data.

Categorial variables in Werner and Svedin (2017) - Fig 2: 1 = No discussion/inaccurate discussion; 2 = Minimal mention; 3 = Accurate but limited discussion; 4 = Thorough discussion.

```{r}
Climate<-read_excel('../Data/RawData/ClimateChange.xlsx', 
                             sheet = "Sheet1", skip = 1)
Climate$`State Hazard Mitigation Plan discussion of Climate Change (Babcock 2013)` <- NULL
Climate$`State Water Agency Practices for Climate Adaptation (EPA 2017)` <- NULL
Climate[,3] <- NULL #weird error in: Climate$`Topic Area of State Practice (EPA 2017)\` <- NULL 

Climate$Notes <- "Categorical: 1-None/Inaccurate disc; 2 - Minimal; 3- Accurate/Limited 4 - Thorough. Source: Werner and Svedin (2017)"

Climate$variable<-"Discussion of Climate Change in plans"
Climate <- merge(Climate, StateList, by.x = "State", by.y = "State_Abbreviation",all.y = TRUE)
colnames(Climate)
Climate$State <- NULL #deletes first instance, which is abbreviations
colnames(Climate)[colnames(Climate) == "State Water Plan discussion of CC (Werner & Svedin 2017)"] <- "value"

Climate <- Climate[c(4,1,3,2)]

#write.csv(Climate,"../Data/FormattedData/ClimateInWaterPlans.csv")
rm(Climate)
```

###Drought plans
Develop proxies for water plan management using drought plans, the most available plan type across the states.

The data was compiled by Rachel McKane.

```{r}
Droughts <- read_excel('../Data/RawData/StateWaterPlans/DroughtPlanSheet.xlsx', 
                             sheet = "Sheet1")

colnames(Droughts)

#keep only select columns
Droughts_subset <- Droughts %>%
  select(State,`Drought Plan (1=Yes)`, `Page Count `, Year)

#cross-check 48 states
Droughts_states <- merge(Droughts_subset, StateList, by = "State") %>%
  select(-State_Abbreviation)

colnames(Droughts_states) <- c("State","DroughtPlan","DroughtPlan_PgCnt","DroughtPlan_Yr")

#write.csv(Droughts_states,"../Data/FormattedData/StateDroughtPlansInfo.csv")
rm(Droughts,Droughts_states,Droughts_subset)
```


###CWNS
Clean Water Needs Survey 2012

```{r}
CWNS<-read.csv("../Data/RawData/CWNS2012.csv")

CWNS <- CWNS %>%
  select(State,Total.ON,Total.UCE,Total.ODN) %>% #these cols represent "Official Needs","Unofficial Cost Estimates","Other Documented Needs" respectively 
  group_by(State) %>%
  summarise_all(sum) %>%
  melt(id = "State") %>%
  group_by(State) %>%
  summarise(TotalNeeds = sum(value))

colnames(CWNS) <- c("State","value")
CWNS$variable <- "Total Infrastructure Needs"
CWNS$Units <- "Dollars"
CWNS$Source <- "EPA Clean Water Needs Survey in 2012"

#subset to 48 contiguous states
CWNS <- merge(CWNS, StateList, by.x = "State", by.y = "State_Abbreviation",all.y = TRUE) #weird enough - no data for SC in the survey...
CWNS$State <- NULL
#repeat columns for SC info
CWNS$variable <- "Total Infrastructure Needs"
CWNS$Units <- "Dollars"
CWNS$Source <- "EPA Clean Water Needs Survey in 2012"

#reorder columns
CWNS <- CWNS[c(5,1,2,3,4)] 

#write.csv(CWNS,"../Data/FormattedData/CleanWaterNeeds.csv")
rm(CWNS)
```

###CWSRF and DWSRF
Federal Grants Received for clean water under revolving fund program and the drinking water state revolving fund since they both have to do with funding associated with water quality 

```{r}
#Should combine both CWSRF and CWSRF into a single variable 
###Drinking Water State Revolving Fund
DWSRF <- read_excel('../Data/RawData/DWSRF.xlsx')

DWSRF_Summarize <- DWSRF %>%
  select(State, `Total Amount `, Units, Variable) %>%
  group_by(State) %>%
  summarize(TotalAmount = sum(`Total Amount `))

DWSRF_Summarize$variable <- "Total Amount (DWSRF)"
DWSRF_Summarize$Year <- "1996-2016"
DWSRF_Summarize$Source <- "DWSRF"

#Onto CWSRF...
CWSRF<-read_excel('../Data/RawData/CWSRF2016.xlsx')

#filter to the 48 states of interest
DWSRF_Summarize$State <- gsub("[[:space:]]*$", "",DWSRF_Summarize$State) #turns out there is a weird space in the states that's affecting the merging so need to remove that; need to only remove the trailing whitespace since some of the states have two word names
DW_QC <- merge(DWSRF_Summarize, StateList, by = "State", all.y = TRUE) #48 states only, all populated!

#explore data within the file
CWSRF$HeaderName <- as.factor(CWSRF$HeaderName)
#levels(CWSRF$HeaderName) #only one of real interest is: "Federal Grants (Dollars)"

CWSRF$RowName <- as.factor(CWSRF$RowName)
#levels(CWSRF$RowName) #only one of real interest is: "*Total Annual Federal Grants"  

CWSRF_Summarize <- CWSRF %>%
  select(-ReportRow) %>%
  filter(HeaderName %in% "Federal Grants (Dollars)") %>%
  filter(RowName %in% "*Total Annual Federal Grants") %>%
  melt(id=c("State","HeaderName","RowName"))

CWSRF_Summarize$variable <- as.numeric(as.character(CWSRF_Summarize$variable)) #silly factor issue

CWSRF_Summarize <- CWSRF_Summarize %>%
  filter(variable > 1995 & variable < 2017) %>% #DWSRF data is limited to these 21 years...
  group_by(State) %>%
  summarize(TotalAmount = sum(value)) #summarize data from 1996 to 2016 into single value

CWSRF_Summarize$variable <- "Cumulative Federal Grants (CWSRF)"
CWSRF_Summarize$Year <- "1996-2016"
CWSRF_Summarize$Source <- "CWSRF"

CombinedFunds <- rbind(CWSRF_Summarize,DWSRF_Summarize)
CombinedFunds$State <- as.factor(CombinedFunds$State)

CombinedFunds <- CombinedFunds %>%
  group_by(State) %>%
  summarise(value = sum(TotalAmount)) 

CombinedFunds_48 <- merge(CombinedFunds, StateList, by = "State", all.y = TRUE) #48 states only, all populated!

CombinedFunds_48$variable <- "Cumulative Federal Grants (CWSRF and DWSRF)"
CombinedFunds_48$Units <- "Dollars"
CombinedFunds_48$Notes <- "CWSRF and DWSRF data between 1996-2016"
CombinedFunds_48$State_Abbreviation <- NULL

#write.csv(CombinedFunds_48,"../Data/FormattedData/FederalGrantsAwarded.csv")
rm(CombinedFunds,CombinedFunds_48,CWSRF,CWSRF_Summarize,DWSRF,DWSRF_Summarize,DW_QC)
```

###Parties in State Legislature
Party in control of state legislature, 1978-2016. Nebraska has unicameral, non-partisan state legislature = NA.

The StateLegParty1990-2017.xls document has number of seats held by each house of the legislature but the website also contains summary info about the leading party in each state that's available in pdfs and has been summarized into the state "PartyControl1978-2016.csv" file

```{r}
Parties<-read.csv("../Data/RawData/PartyControl1978-2016.csv")

Parties_48 <- merge(Parties, StateList, by = "State", all.y = TRUE) 

Parties_48 <- Parties_48 %>%
  melt(id="State") 

Parties_48$value[is.na(Parties_48$value)] <- "NP" #less troublesome of a label :-)
Parties_48$variable <- gsub("X","",Parties_48$variable) #remove leading "X" marker
Parties_48$variable <- as.numeric(Parties_48$variable)

Parties_48_count <- Parties_48 %>%
  filter(variable > 1998) %>% #over last 10 state elections
  group_by(State,value) %>%
  summarise(counter = n()) %>%
  #dcast(State ~ value, value.var = "counter")
  filter(counter == max(counter)) %>%
  mutate(Percentage = counter/10*100) 

#Indiana has a 5-5 Rep-Split while Nevada has a 4-4 Dem-Split  so will just keep the dominant party for these 2 states
Parties_48_count %>%
  filter(State %in% c("Indiana","Nevada")) %>%

Parties_48_final <- Parties_48_count[!(Parties_48_count$State=="Indiana" & Parties_48_count$value=="Split"),]
Parties_48_final <- Parties_48_final[!(Parties_48_final$State=="Nevada" & Parties_48_final$value=="Split"),]

#update column names
colnames(Parties_48_final)[colnames(Parties_48_final) == "value"] <- "State Leading Party"
colnames(Parties_48_final)[colnames(Parties_48_final) == "Percentage"] <- "Percentage_State Leading Party"

Parties_48_final <- Parties_48_final %>%
  select(-counter) %>%
  melt(id="State")

Parties_48_final$Notes <- "State Legislature power between 2000-2016"

#write.csv(Parties_48_final,"../Data/FormattedData/StateLegislatureParties.csv")
rm(Parties, Parties_48,Parties_48_count, Parties_48_final)

# #Brief exploration of the xls document...
# Parties_Upper <- read_excel('../Data/RawData/StateLegParty1990-2017.xls', sheet = "Upper", skip = 2) #note: data is of Aug 2017
# #replicate year info across the rows
# 
# Parties_upper_t <- as.data.frame(t(Parties_Upper))
# Parties_upper_t <- rownames_to_column(Parties_upper_t, var= "Year")
# 
# while(length(ind <- which(Parties_upper_t$Year == "")) > 0){
#   Parties_upper_t$Year[ind] <- Parties_upper_t$Year[ind -1]
# }
# 
# Parties_Upper_Yrs <- as.data.frame(t(Parties_upper_t))
# colnames(Parties_Upper_Yrs) <- paste(as.character(unlist(Parties_Upper_Yrs[1,])),as.character(unlist(Parties_Upper_Yrs[2,])),sep="_")
# Parties_Upper_Yrs<- Parties_Upper_Yrs[-2,] #delete states row
# Parties_Upper_Yrs<- Parties_Upper_Yrs[-1,] #delete years row
# 
# colnames(Parties_Upper_Yrs)[1] <- "State"
# 
# Parties_Upper_Yrs$State <- gsub("\\*","",Parties_Upper_Yrs$State) #get rid of weird asterisks
# 
# #filter for the 48 states
# Parties_Upper_Yrs_48 <- merge(Parties_Upper_Yrs, StateList, by = "State", all.y = TRUE) #48 states only, all populated!
# rm(Parties, Parties_Upper, Parties_upper_t, Parties_Upper_Yrs, Parties_Upper_Yrs_48)
```

###Federal voting margin
per JG's earth future's analysis...

averaged over the 2008 and 2012 elections. "partisan political leanings of states and MSA's by the Cook Partisan Voting Index (PVI) Wasserman [2013]. The index measures the difference between 70 the percentage of the two-party vote received by Democratic presidential candidates in a city or MSA and the percentage received in the national election. Positive or negative values indictate a city or state leans Democratic or Republican, respectively, relative to the national average."

Findings." States with more positive (Democratic-leaning) PVI have greater propensity to adopt conservation policies, as do MSAs whose PVI is greater (more likely to vote Democratic) than the rest of the state.. TODO: discuss robustness with respect to consistency about state-level aridity and PVI across all models. What emerges in the big picture is that cities in states with higher (more Democratic-leaning) PVI's and more arid climates tend to adopt more conservation measures, including more requirements and more rebates"

(Positive PVI indicates a greater Democratic vote share than the national average and negative PVI a greater Republican vote share)


```{r}
PVI <- read_csv('../Data/RawData/pvi_by_state.csv')

#filter for 48 states
PVI_48 <- merge(PVI, StateList, by.x = "state.abb", by.y = "State_Abbreviation", all.y = TRUE)

PVI_48 <- PVI_48 %>%
  select(State, pvi)

colnames(PVI_48)[colnames(PVI_48) == "pvi"] <- "value"
PVI_48$variable <- "Cook's Partisan Voting Index"
PVI_48$Units <- "Index"
PVI_48$Year <- "2008 and 2012 elections"
PVI_48$Notes <- "Positive values indicate greater Dem vote rel. to natioanl avg (vv for negative PVI and Rep votes). Source: JG's Earth Future Paper"

#write.csv(PVI_48,"../Data/FormattedData/PVI.csv")
rm(PVI, PVI_48)
```

###Sewage Disposal
all homes 1970, 1980, 1990 is from the census while we have limited information about new homes built 2009-2016. For now, just going to process rate of homes in each state on public sewer as of 1990

```{r}
Sewage<-read_excel('../Data/RawData/SewageDisposal.xlsx', sheet = "SewageDisposal70_90", col_names = FALSE)

#fix colnames
colnames(Sewage) <- paste(as.character(unlist(Sewage[1,])),as.character(unlist(Sewage[2,])),sep="_")
Sewage<- Sewage[-2,] #delete #/% row
Sewage<- Sewage[-1,] #delete type of disposal row

Sewage$NA_Year <- as.numeric(Sewage$NA_Year)

Sewage_final <- Sewage %>%
  filter(NA_Year == 1990) %>% #only keep 1990 data
  select(NA_State,NA_Year,`Public Sewer_Percent`)

colnames(Sewage_final) <- c("State","Year","value")
Sewage_final$value <- as.numeric(Sewage_final$value)
Sewage_final$value <- Sewage_final$value*100 #so values are actual %s

#filter to continguous 48 states
Sewage_final_48 <- merge(Sewage_final, StateList, by.x = "State", by.y = "State_Abbreviation", all.y = TRUE) #48 states only, all populated!
Sewage_final_48$State <- NULL #deletes first instance of abbrevations
Sewage_final_48 <- Sewage_final_48[c(3,2,1)]#reorder columns

#metadata columns
Sewage_final_48$Units <- "Percentage"
Sewage_final_48$variable <- "Populace on Public Sewer"
Sewage_final_48$Notes <- "Source: US Census"

#write.csv(Sewage_final_48,"../Data/FormattedData/SewageDisposal1990.csv")
rm(Sewage,Sewage_final,Sewage_final_48)
```

###Affordability
Mack's research that highlights the % of state tracts in each state with high risk for drinking water affordability issues

```{r}
Affordability <-read_excel('../Data/RawData/WaterBill.xlsx', sheet = "waterbill")
colnames(Affordability)

Affordability$value <- 100 * (Affordability$`Percent of State's Tracts AtRisk` + Affordability$`Percent of State's Tracts HighRisk`)

Affordability <- Affordability %>%
  select(State,value)

#Ensure all 48 contiguous state are on the list
Affordability_final <- merge(Affordability, StateList, by = "State", all.y = TRUE) #48 states only, all populated!
Affordability_final$State_Abbreviation <- NULL 

Affordability_final$Units <- "Percentage"
Affordability_final$variable <- "State Tracts at either High Risk or At-Risk for water poverty or unaffordable water svs"
Affordability_final$Year <- 2015
Affordability_final$Source <- "Mack and Wrase (2017)"

#write.csv(Affordability_final,"../Data/FormattedData/WaterPovertyRisk.csv")
rm(Affordability,Affordability_final)
```

###Public Supply 
Proportion of state populace on public supply for drinking water, 1960-2010
Issues with provenance still need to be resolved with Brandi (i.e., what the variables are measuring exactly)

```{r}
PublicSupply<-read.csv("../Data/RawData/PublicSupplyByState1960_2010.csv")

#rename columns 
colnames(PublicSupply)[colnames(PublicSupply) == "ï.."] <- "State"
colnames(PublicSupply)[colnames(PublicSupply) == "X2010"] <- "value"

PublicSupply <- PublicSupply %>%
  select(State, value)

PublicSupply_QC <- merge(PublicSupply, StateList, by = "State", all.y = TRUE) #48 states only, all populated!

PublicSupply$value <- PublicSupply$value * 100 #convert to actual percents

PublicSupply$Units <- "Percentage"
PublicSupply$Year <- 2010
PublicSupply$variable <- "Populace on public supply for drinking water"
PublicSupply$Notes <- "Source: USGS"

#write.csv(PublicSupply,"../Data/FormattedData/PublicSupply_DW.csv")
rm(PublicSupply, PublicSupply_QC)
```

###Waterborne Illnesses 
from drinking water, recreational waters, and environmental or undetermined waters
FY2011-2012. Other years (to 1971) available at https://www.cdc.gov/healthywater/surveillance/index.html
```{r}
#Illnesses by Drinking Water
DW_Illness <-read_excel('../Data/RawData/outbreaks_dw20112012.xlsx', skip = 1)
DW_Illness <- DW_Illness[,-c(14,15)] #weird empty trailing columns

#rename columns of interest
colnames(DW_Illness)[colnames(DW_Illness) == "State/ Jurisdiction "] <- "State"
colnames(DW_Illness)[colnames(DW_Illness) == "No. cases "] <- "value"

DW_Illness <- DW_Illness %>%
  select(State, value) %>%
  group_by(State) %>%
  summarise(value = sum(value))

DW_Illness$State <- gsub("[[:space:]]*$", "",DW_Illness$State) #fix trailing spaces after state names
DW_Illness_48 <- merge(DW_Illness, StateList, by = "State", all.y = TRUE)
DW_Illness_48$variable <- "DW"

#Illnesses by Recreational Water
Rec_Illness <-read_excel('../Data/RawData/outbreaks_rw20112012.xlsx', skip = 1)

colnames(Rec_Illness)[colnames(Rec_Illness) == "Jurisdiction "] <- "State"
colnames(Rec_Illness)[colnames(Rec_Illness) == "No. cases§ "] <- "value"

Rec_Illness <- Rec_Illness %>%
  select(State, value) %>%
  group_by(State) %>%
  summarise(value = sum(value))

Rec_Illness$State <- gsub("[[:space:]]*$", "",Rec_Illness$State) #fix trailing spaces after state names
Rec_Illness_48 <- merge(Rec_Illness, StateList, by = "State", all.y = TRUE)
Rec_Illness_48$variable <- "Rec"

#Illnesses by environmental water
Env_Illness <-read_excel('../Data/RawData/outbreaks_euw20112012.xlsx', skip = 1)

colnames(Env_Illness)[colnames(Env_Illness) == "State/Jurisdiction "] <- "State"
colnames(Env_Illness)[colnames(Env_Illness) == "No. cases "] <- "value"

Env_Illness <- Env_Illness %>%
  select(State, value) %>%
  group_by(State) %>%
  summarise(value = sum(value))

Env_Illness$State <- gsub("[[:space:]]*$", "",Env_Illness$State) #fix trailing spaces after state names
Env_Illness_48 <- merge(Env_Illness, StateList, by = "State", all.y = TRUE)
Env_Illness_48$variable <- "Env"

#All Illnesses together
waterillness<-rbind(DW_Illness_48,Rec_Illness_48,Env_Illness_48)
waterillness$value[is.na(waterillness$value)] <- 0 #replace NAs with 0s
waterillness$State_Abbreviation <- NULL

waterillness_sum <- waterillness %>%
  group_by(State) %>%
  summarise(value = sum(value))

waterillness_sum$variable <- "Water illnesses related to DW, Rec, and Env"
waterillness_sum$Units <- "Number of Cases"
waterillness_sum$Year <- "2011-2012"
waterillness_sum$Notes <- "Source: CDC"

#write.csv(waterillness_sum,"../Data/FormattedData/CDCWaterIllnesses.csv")
rm(Env_Illness,Env_Illness_48,DW_Illness,DW_Illness_48,Rec_Illness,Rec_Illness_48,waterillness,waterillness_sum)
```

###Dams
NID storage measured in Acre-Feet
```{r}
#Dams with water supply as primary purpose
Supply<-read_excel("../Data/RawData/DamsWaterSupplyAllPurposes.xlsx")

colnames(Supply)

Supply <- Supply %>%
  select(State, `Owner Type`, NIDID, `NID Storage`, `NID Height (Ft.)`, `Primary Purpose`) #only keep most pertinent data

Supply <- Supply[complete.cases(Supply), ] #get rid of rows with NAs; only 1 that this applies to

#Check to see what dams are in the hydroelectric dataset but not in the water supply dataset, esp. since the WaterSupply one has other uses as the primary purpose
Hydro<-read_excel("../Data/RawData/DamsHydroelectricAllPurposes.xlsx")

colnames(Hydro)

Hydro <- Hydro %>%
    select(State, `Owner Type`, NIDID, `NID Storage`, `NID Height (Ft.)`, `Primary Purpose`) #only keep most pertinent data

#combine both dam uses 
Dams_comparison <- merge(Supply, Hydro, by=c("NIDID","Owner Type","State","Primary Purpose","NID Storage","NID Height (Ft.)"), all = TRUE) #looks there were 372 dams that overlapped
Dams_comparison$`Owner Type`[Dams_comparison$`Owner Type`=="-"]<-"Not Listed"

Dams_Owner <- Dams_comparison %>%
  select(State, `Owner Type`, `NID Storage`) %>%
  group_by(State, `Owner Type`) %>%
  summarise(Storage_AF = sum(`NID Storage`)) %>%
  group_by(State) %>%
  mutate(Total_AF = sum(Storage_AF)) %>%
  mutate(Storage_Percent = Storage_AF/Total_AF * 100) %>% #% of each state's dams in different authorities
  filter(Storage_Percent == max(Storage_Percent)) %>%
  select(-Storage_AF)
  #spread(`Owner Type`, Storage_Percent)

colnames(Dams_Owner)[colnames(Dams_Owner) == "Owner Type"] <- "Leading Dam Owner"
colnames(Dams_Owner)[colnames(Dams_Owner) == "Storage_Percent"] <- "Percentage_Leading Dam Owner"

#only keep contiguous 48
Dams_48 <- merge(Dams_Owner, StateList, by.x="State", by.y="State_Abbreviation", all.y = TRUE)
Dams_48$State <- NULL
Dams_48 <- Dams_48[c(4,2,1,3)]

Dams_48_Total <- Dams_48 %>%
  select(State,Total_AF) %>%
  melt(id="State")

Dams_48_Total$variable <- "Total Dam Storage"
Dams_48_Total$Units <- "Acre-Ft"
Dams_48_Total$Source <- "USACE NID as of 2015"

Dams_48_Owners <- Dams_48 %>%
  select(-Total_AF) %>%
  melt(id = "State")

Dams_48_Owners$Source <- "USACE NID as of 2015"

#write.csv(Dams_48_Total, "../Data/FormattedData/TotalDamStorage.csv")
#write.csv(Dams_48_Owners,"../Data/FormattedData/LeadingDamOwnership.csv")
rm(Dams_48,Dams_48_Owners,Dams_48_Total,Dams_comparison,Dams_Owner,Hydro,Supply)
```


###State Withdrawals and State uses
The original source file is https://water.usgs.gov/watuse/data/2010/USGSCircular1405-tables1-14.xlsx but I had already processed some of this data for the proposal. So in the interest of time, I will continue to work off the intermediate file. Note: the original file is saved in the RawData folder in case it's needed for verification.

Note: raw values in MGDs

```{r}
WaterUse<-read_excel("../Data/RawData/State_MonthlyPrecip_Temp_WaterUse_ImpairedStreams.xlsx", sheet = "StateUse_MGD")

#quick QC of 48 cont. states
WaterUse_QC <- merge(WaterUse, StateList, by = "State", all.y = TRUE) #looks good :-) - all 48 states present and populated
rm(WaterUse_QC)

##----Total water use in the state 
WaterUse_Total <- WaterUse %>%
  select(State,Total) %>%
  rename(value = Total) %>%
  mutate(variable = "Total water use") %>%
  mutate(Units = "Million gallons per day") %>%
  mutate(Year = 2010) %>%
  mutate(Notes = "Source: USGS Circular")

#write.csv(WaterUse_Total, "../Data/FormattedData/TotalWaterUse.csv")
rm(WaterUse_Total)

##---Water use by percentage: sw vs gw
WaterUse_Source <- WaterUse %>%
  select(State, matches("GW|SW"), matches("Total")) %>% #contains only total gw and sw uses

WaterUse_Source_TotalPercents <- WaterUse_Source %>%
  mutate(Total = GW_Total + SW_Total) %>%
  mutate(value = SW_Total / Total * 100) %>% #
  select(State, value)

WaterUse_Source_TotalPercents$variable <- "Surface water use relative to total water use"
WaterUse_Source_TotalPercents$Units <- "Percentage"
WaterUse_Source_TotalPercents$Year <- 2010
WaterUse_Source_TotalPercents$Notes <- "Source: USGS Circular"

#write.csv(WaterUse_Source_TotalPercents, "../Data/FormattedData/WaterUsebySource_SW.csv")
rm(WaterUse_Source_TotalPercents, WaterUse_Source)

##---water use by fresh vs saline across sw and gw
WaterUse_Source_Saline <- WaterUse %>%
  select(State, matches("GW|SW"), -matches("Fresh")) %>% #contains both saline and freshwater information by source
  mutate(Total_Saline = GW_Saline + SW_Saline, Total_Total = GW_Total + SW_Total) %>%
  melt(id = "State")

#create better ID'ing variables
WaterUse_Source_Saline$SplitString <- str_split(WaterUse_Source_Saline$variable, "_")
WaterUse_Source_Saline$Source <- sapply(WaterUse_Source_Saline$SplitString,"[[",1) #sapply to get 1st element of gw vs sw
WaterUse_Source_Saline$Quality <- sapply(WaterUse_Source_Saline$SplitString,"[[",2) #sapply to get 2nd element of saline vs total
WaterUse_Source_Saline$SplitString <- NULL

#calculate percentage of saline water use by source
WaterUse_Source_Saline_Percents <- WaterUse_Source_Saline %>%
  group_by(State, Source) %>%
  mutate(SalinePercent = round(value/value[Quality == "Total"] * 100, 2)) %>%
  filter(Quality %in% "Saline") %>%
  select(State, SalinePercent, Source) %>%
  as.data.frame()

colnames(WaterUse_Source_Saline_Percents)[colnames(WaterUse_Source_Saline_Percents) == "SalinePercent"] <- "value"
WaterUse_Source_Saline_Percents$variable <- "Saline water use relative to total water use by source"  
WaterUse_Source_Saline_Percents$Units <- "Percentage"
WaterUse_Source_Saline_Percents$Year <- 2010
WaterUse_Source_Saline_Percents$Notes <- ".GW = groundwater, SW= surface water. Source: USGS Circular"

#write.csv(WaterUse_Source_Saline_Percents, "../Data/FormattedData/WaterUsebySource_Quality.csv")
rm(WaterUse_Source_Saline_Percents, WaterUse_Source_Saline)

##----Water use by sector
WaterUse_Sector <- WaterUse %>%
  select(-matches("GW|SW")) %>%
  melt(id="State") %>%
  group_by(State) %>%
  mutate(Percentage = round(value/value[variable == "Total"] * 100,2)) %>% #round to 2 decimal pnts
  filter(!variable %in% "Total") %>%
  select(-value) %>% #don't need the raw MGD #s
  rename(value = Percentage) %>% #rename percentage column to value
  rename(Sector = variable) %>%
  mutate(variable = "Proportion of water use by sector") %>%
  mutate(Units = "Percentage") %>%
  mutate(Year = 2010) %>%
  mutate(Notes = "Source: USGS Circular")

#write.csv(WaterUse_Sector,"../Data/FormattedData/WaterUse_Sector.csv")
rm(WaterUse, WaterUse_Sector)
```

###Electricity Generation
Electricity generation by state for all fuels (utility-scale). 

Also have data for Conventional Hydro and Pumped Hydro if that becomes important

```{r}


#net generation for all fuels, utility-scale in MGW
Allfuels<-read.csv("../Data/RawData/NetGenAllFuelsByState2001_2016.csv",skip=4,na.strings = c("", "NA","--"))

Allfuels_Clean <- Allfuels %>%
  select(description, units, X2016) %>%
  filter(complete.cases(.)) %>%
  mutate(SplitString = str_split(description, " : ")) %>%
  mutate(State = sapply(SplitString, "[[" , 1)) %>%
  mutate(variable = sapply(SplitString, "[[" , 2)) %>%
  filter(variable %in% "all sectors") %>%
  rename(value = X2016) %>%
  rename(Units = units) %>%
  select(-SplitString, -description) %>%
  mutate(Year = 2016) %>%
  mutate(Notes = "Source: EIA. Producers include all commercial, industrial, electric power, electric utility, and independent power producers")

Allfuels_Clean$variable <- "Net generation by all fuels and all utility-scale producers"
Allfuels_Clean <- Allfuels_Clean[c(3,2,4,1,5,6)]

#quick QC to ensure 48 contiguous are all populated
Allfuels_Clean_QC <- merge(Allfuels_Clean, StateList, by="State", all.y = TRUE) #all good :-)

#write.csv(Allfuels_Clean,"../Data/FormattedData/NetElectricityGeneration.csv")

##---Hydropower as percentage of state's net generation 
Conv_Hydro<-read.csv("../Data/RawData/NetGenConventionalHydroByState2001_2016.csv",skip=4,na.strings = c("", "NA","--","NM"))
Pump_Hydro<-read.csv("../Data/RawData/NetGenPumpedHydroByState2001_2016.csv",skip=4,na.strings = c("", "NA","--","NM"))

AllHydro <- rbind(Conv_Hydro,Pump_Hydro)

Hydro_Clean <- AllHydro %>%
  select(description, units, X2016) %>%
  filter(complete.cases(.)) %>%
  mutate(SplitString = str_split(description, " : ")) %>%
  mutate(State = sapply(SplitString, "[[" , 1)) %>%
  mutate(variable = sapply(SplitString, "[[" , 2)) %>%
  filter(variable %in% "all sectors") %>%
  rename(value1 = X2016) %>%
  mutate(value = as.numeric(as.character(value1))) %>%
  group_by(State) %>%
  summarise(value = sum(value)) %>%
  mutate(variable = "Net hydropower generation") %>%
  mutate(Units = "thousand megawatthours") %>%
  mutate(Year = 2016) %>%
  mutate(Notes = "Source: EIA. Includes both conventional and pumped hydropower generation")

#make sure all 48 states are in Hydro
Hydro_Clean_48 <- merge(Hydro_Clean, StateList, by = "State", all.y = TRUE)

Hydro_Clean_48 <- Hydro_Clean_48 %>%
  select(-State_Abbreviation) %>%
  mutate(variable = "Net hydropower generation") %>%
  mutate(Units = "thousand megawatthours") %>%
  mutate(Year = 2016) %>%
  mutate(Notes = "Source: EIA. Includes both conventional and pumped hydropower generation") %>%
  mutate(value = ifelse(is.na(value), 0, value)) #replace NAs with 0s

Hydro_vs_Total <- rbind(Hydro_Clean_48,Allfuels_Clean)

Hydro_Clean_Percent <- Hydro_vs_Total %>%
  mutate(value = as.numeric(value)) %>%
  group_by(State) %>%
  mutate(Percentage = round(value/value[variable == "Net generation by all fuels and all utility-scale producers"] * 100, 2)) %>%
  filter(variable %in% "Net hydropower generation") %>%
  select(-value) %>% #get rid of hydropower in mwh
  rename(value = Percentage) %>% #move value column to percentage
  mutate(Units = "Percentage")

#reorder columns for consistency
Hydro_Clean_Percent <- Hydro_Clean_Percent[c(1,6,2,3,4,5)]

#write.csv(Hydro_Clean_Percent,"../Data/FormattedData/PercentageofHydropowerGeneration.csv")
rm(Allfuels, Allfuels_Clean, Allfuels_Clean_QC, AllHydro, Conv_Hydro, Hydro_Clean, Hydro_Clean_48, Hydro_Clean_Percent, Hydro_vs_Total, Pump_Hydro)
```

### Aridity Index
Following JG's analysis, there is strong correlation between precipitation and temperature datasets so I will construct the Koppen aridity index, where AI = P/(T+33), with P in mm and T in C. Large values indicate wetter conditions and smaller values to drier conditions. 

Monthly precipitation (in inches) 01/1895-12/2016
Monthly average temperature (in degrees Farenheit) 01/1895-12/2016

```{r}
Precip<-read_excel("../Data/RawData/State_MonthlyPrecip_Temp_WaterUse_ImpairedStreams.xlsx", sheet = "Precip_Inches")
Temp<-read_excel("../Data/RawData/State_MonthlyPrecip_Temp_WaterUse_ImpairedStreams.xlsx", sheet = "Temp_F")

Precip_formatted <- Precip %>%
  filter(! YrMo > 201612) %>% #subset out the four months of 2017 data
  melt(id="YrMo") %>% 
  rename(State = variable) %>%
  mutate(value = value * 25.4) %>% #convert inches to mm
  mutate(Year = substr(as.character(YrMo),1,4)) %>%
  mutate(Month = substr(as.character(YrMo),5,6)) %>%
  mutate(variable = "Precip_mm")

Temp_formatted <- Temp %>%
  filter(! YrMo > 201612) %>% #subset out the four months of 2017 data
  melt(id="YrMo") %>% 
  rename(State = variable) %>%
  mutate(value = (value - 32) * 5 / 9) %>% #convert deg F to deg C
  mutate(Year = substr(as.character(YrMo),1,4)) %>%
  mutate(Month = substr(as.character(YrMo),5,6)) %>%
  mutate(variable = "Temp_C")

Met_data <- rbind(Precip_formatted, Temp_formatted)

Met_data <- Met_data %>%
  spread(variable, value) %>%
  mutate(value = Precip_mm/(Temp_C + 33)) %>%
  mutate(variable = "Koppen Aridity Index") %>%
  mutate(Units = "Index") %>%
  mutate(Notes = "Large value indicate wet conditions and v.v. Following JG's analysis, Koppen formula = P_mm/(T_C+33)") %>%
  select(-Precip_mm, -Temp_C)
#write.csv(Met_data,"../Data/FormattedData/AridityIndex_TimeSeries.csv")

Met_data_annualavg_median <- Met_data %>%
  group_by(State,Year) %>%
  summarise(Annual_AI = mean(value)) %>% #calculate annual means
  group_by(State) %>%
  summarise(Median_AI = median(Annual_AI), Stdev_Interannual_AI = sd(Annual_AI)) %>% #calculate median of annual means
  melt(id = "State") %>%
  mutate(variable = ifelse(variable == "Median_AI", "Median of Annual Aridity Index", "Interannual Variability (Stdev of annual means)")) %>%
  mutate(Year = "Jan 1895 - Dec 2016") %>%
  mutate(Units = "Index") %>%
  mutate(Notes = "Large value indicate wet conditions and v.v. Following JG's analysis, Koppen formula = P_mm/(T_C+33). Source: NOAA for met data.")

Met_data_intraannual <- Met_data %>%
  group_by(State,Month) %>%
  summarize(MonthlyMeans_AI = mean(value)) %>%
  group_by(State) %>%
  summarize(Intraannual = sd(MonthlyMeans_AI)) %>%
  melt(id = "State") %>%
  mutate(variable = "Intraannual Variability (Stdev of monthly means)") %>%
  mutate(Year = "Jan 1895 - Dec 2016") %>%
  mutate(Units = "Index") %>%
  mutate(Notes = "Large value indicate wet conditions and v.v. Following JG's analysis, Koppen formula = P_mm/(T_C+33). Source: NOAA for met data.")
  
AI_Stat_data <- rbind(Met_data_annualavg_median, Met_data_intraannual)

#write.csv(AI_Stat_data,"../Data/FormattedData/AridityIndex_Median_Variability.csv")
rm(Met_data,Met_data_annualavg_median,Met_data_intraannual,Precip,Precip_formatted,Temp, Temp_formatted, AI_Stat_data)
```

### Misc Variables from Water Conservation analysis
In the interest of time, going to pull some extra variables from Hornberger et al (2015) WRR paper on variable correlations to a state's conservation practices. The raw data file is from GMH.

```{r}
Misc_data <- read_excel("../Data/RawData/data_48statesContiguous.xlsx", sheet = "awe_scores")

colnames(Misc_data)

Misc_data_select <- Misc_data %>%
  select(state,score,kwh) %>%
  filter(complete.cases(.)) %>% #get rid of weird trailing NAs
  melt(id="state") %>%
  rename(State = state) %>%
  mutate(Units = ifelse(variable == "kwh","Kilowatthours", "Index")) %>%
  mutate(Year = ifelse(variable == "kwh",2010, 2011)) %>%
  mutate(variable = ifelse(variable == "kwh","Annual electricity use per capita","AWE Conservation Score")) %>%
  mutate(Notes = "Source: Hornberger et al (2015) WRR, Table 1")
  
#write.csv(Misc_data_select,"../Data/FormattedData/ConservationScore_ElectricityUse.csv")
rm(Misc_data,Misc_data_select)
```

###EPA Regions
EPA Regions that each state belongs to

```{r}
EPARegions <- read_csv("../Data/RawData/EPA Regions.csv")

#only keep 48 contiguous 48 states
EPA_48 <- merge(EPARegions, StateList, by = "State_Abbreviation", all.y = TRUE) #all 48 there and populated :-)

#Clean up df
EPA_Final <- EPA_48 %>%
  select(-State_Abbreviation) %>%
  rename(value = `EPA Region`) %>%
  mutate(variable = "EPA Regions") %>%
  mutate(Units = "Region") 

EPA_Final <- EPA_Final[c(3,1,4,5,2)]

#write.csv(EPA_Final, "../Data/FormattedData/EPARegions.csv")
rm(EPARegions,EPA_Final,EPA_48)
```

###Land Use
Percentage of each state in the major land use categories. Original Units are 1000 acres

```{r}
LandUses <- read_excel("../Data/RawData/Summary_Table_1_major_uses_of_land_by_region_and_state_2012.xls", skip = 2)

#keep complete cases
LandUses_subset <- LandUses %>%
  filter(complete.cases(.)) %>%
  select(`Regions and States`, `Cropland  5/`, `Grassland pasture and range  5/`, `Forest-use land`, `Urban areas `, `Total land area 2/`) %>%
  rename(State = `Regions and States`) %>% #rename columns into reasonable titles
  rename(Cropland = `Cropland  5/`) %>%
  rename(Grassland = `Grassland pasture and range  5/`) %>%
  rename(Forest = `Forest-use land`) %>%
  rename(Urban = `Urban areas `) %>%
  rename(Total = `Total land area 2/`) %>%
  mutate(Cropland = as.numeric(Cropland)) %>% #format columns into numeric
  mutate(Grassland = as.numeric(Grassland)) %>%
  mutate(Forest = as.numeric(Forest)) %>%
  mutate(State = str_trim(State, side = "both")) #trim weird trailing white spaces

#only keep 48 contiguous 48 states
LandUses_48 <- merge(LandUses_subset, StateList, by = "State", all.y = TRUE) #all 48 there and populated :-)

#Create final df
LandUses_48_final <- LandUses_48 %>%
  select(-State_Abbreviation) %>%
  melt(id = "State") %>%
  group_by(State) %>%
  mutate(value = value/value[variable == "Total"] * 100) %>%
  filter(!variable %in% "Total") %>%
  mutate(variable = paste("Land Use",variable, sep =": ")) %>%
  mutate(Units = "Percentage") %>%
  mutate(Years = 2012) %>%
  mutate(Notes = "Source: USDA ERS")

#write.csv(LandUses_48_final, "../Data/FormattedData/LandUse.csv")
rm(LandUses,LandUses_subset,LandUses_48, LandUses_48_final)
```

###Gross Domestic Product and Price Parity Index
Gross domestic product is for all industries in the state

"Regional Price Parities (RPPs) measure the differences in price levels across states and metropolitan areas for a given year and are expressed as a percentage of the overall national price level."

Full web-links of the original data sources are in the excel document. PDFs of the screen shots of the website are also saved in the raw data folder if needed.

```{r}
GDP_RPP <- read_excel("../Data/RawData/State GSP&RPP - 2010.xlsx")

colnames(GDP_RPP) <- c("State","Gross Domestic Product","Regional Price Parities","S","N")

#only keep 48 contiguous 48 states
GDP_RPP_48 <- merge(GDP_RPP, StateList, by.x = "State", by.y = "State_Abbreviation" ,all.y = TRUE) #all 48 there and populated :-)
GDP_RPP_48$State <- NULL

GDP_RPP_48_Final <- GDP_RPP_48 %>%
  select(State, `Gross Domestic Product`, `Regional Price Parities`) %>%
  melt(id = "State") %>%
  mutate(value = ifelse(variable == "Gross Domestic Product", value * 1000000, value)) %>% #convert GSP units from millions of dollars to dollars
  mutate(Units = ifelse(variable == "Gross Domestic Product", "Dollars (Current value)", "Index")) %>% 
  mutate(Year = 2010) %>%
  mutate(Notes = "Source: US Bureau of Economic Analysis")

#write.csv(GDP_RPP_48_Final, "../Data/FormattedData/GDP_RPP.csv")
rm(GDP_RPP,GDP_RPP_48,GDP_RPP_48_Final)
```