---
title: "Water Quality: Master Data Frame Variable Exploration"
author: "TG"
date: "November 6, 2017"
output: pdf_document
---

## Background
This Rmd file captures the code used to generate a master dataframe to faciliate follow-up exploratory data analysis of water quality issues in the states 

####Set-up environment
```{r setup, include=FALSE}
rm(list=ls()) #clears variables in global environment :-)
#dev.off() #clears plots in history

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "//vu1file.it.vanderbilt.edu/VIEE/Shared/Water Security/Water Quality")

#setwd('//vu1file.it.vanderbilt.edu/VIEE/Shared/FEW/RCode') #Use directory E when on desktop and G on laptop
getwd()
```
####Import Packages
```{r message=FALSE} 
library(tidyverse) #includes dplyr, ggplot2, tidyr, purr, readr, and tibble
library(reshape2)
library(cowplot)
#library(viridis)
#library(stringr)

#StateList <- read.csv('../Data/RawData/StateList.csv')
#StateList$State <- as.character(StateList$State)
```

####Import all of the data
```{r, echo=FALSE, warning=FALSE}
#create list of filenames and filepaths
filepaths <- paste("../Data/FormattedData/", list.files("../Data/FormattedData"), sep="") 
filenames <- gsub("\\.csv$","", list.files("../Data/FormattedData", pattern="\\.csv$"))

#read in the csvs and assign to separate dfs
for(i in 1:length(filenames)){
  assign(filenames[i], read.csv(filepaths[i])) #read in as regular dfs; read_csv = tbl df
}

rm(i, filepaths, filenames)
```

###Generate DFs
-generate master response variable df with the most important impaired data as well as the NCDC and EPA region data

-generate master predictor variable df with the most important data (note: will not include illnesses)

```{r}
rm(CDCWaterIllnesses) #not going to include illnesses in the analysis due to framework approach

#----------Construct df of misc state static info, including normalization variables into single df
ClimRegions <- dcast(ClimRegions, State ~ variable)

levels(StateAreas$variable)
StateAreas <- StateAreas %>%
  filter(variable %in% c("Total Area_SqMi","Land Area_SqMi")) %>%
  dcast(State ~ variable)

StateLatLong <- dcast(StateLatLong, State ~ variable)

EPARegions <- dcast(EPARegions, State ~ variable)

levels(PopulationSize_Growth$variable)
PopulationSize <- PopulationSize_Growth %>%
  filter(variable %in% "Population Size") %>%
  dcast(State ~ variable)

StateInfo <- do.call(cbind, list(StateAreas,StateLatLong,ClimRegions,EPARegions, PopulationSize))
StateInfo <- StateInfo[!duplicated(as.list(StateInfo))] #drop duplicate state label columns
rm(ClimRegions,StateAreas,StateLatLong,PopulationSize,EPARegions) #remove original datafiles

#---------------------Construct response matrix of impaired data
levels(StateImpairedWaters$variable)
Responsedf <- StateImpairedWaters %>%
  filter(variable %in% c("ImpairedLakes_Percent","ImpairedStreams_Percent")) %>%
  dcast(State ~ variable)
rm(StateImpairedWaters)

#------------------Construct predictor matrix of df
rm(AridityIndex_TimeSeries) #no need for this at this point

#Exploration and selection of the variables
levels(Age$variable) #keep all 3 of thse for now: 20 to 34, 62 and over, Median age; % or years
Age <- dcast(Age, State ~ variable)
colnames(Age) <- c("State","Age:20 to 34_percent","Age:62+_percent","Median Age_years")

levels(AridityIndex_Median_Variability$variable) #keep all 3: median, interannual var, and intraannual var
AridityIndex_Median_Variability <- dcast(AridityIndex_Median_Variability, State ~ variable)

levels(CleanWaterNeeds$variable) #units: dollars
CleanWaterNeeds <- dcast(CleanWaterNeeds, State ~ variable)
colnames(CleanWaterNeeds) <- c("State","CleanWaterNeeds_Dollars")

levels(ClimateInWaterPlans$variable) #units: categorical
ClimateInWaterPlans <- dcast(ClimateInWaterPlans, State ~ variable)
colnames(ClimateInWaterPlans) <- c("State","CliminWaterPlans_0poorto4strong")

levels(ConservationScore_ElectricityUse$variable) #units: index and Kwh
ConservationScore_ElectricityUse <- dcast(ConservationScore_ElectricityUse, State ~ variable)
colnames(ConservationScore_ElectricityUse) <- c("State","AnnualElecUseperCapita_KwH","AWE Conservation Score")

levels(Education$variable) #units: %
Education <- dcast(Education, State ~ variable)
colnames(Education) <- c("State","Education: Bachelor's or higher_percent","Education: Graduate or Professional degree_percent","Education: High School or higher_percent")

levels(FederalGrantsAwarded$variable)
FederalGrantsAwarded <- dcast(FederalGrantsAwarded, State ~ variable)
colnames(FederalGrantsAwarded) <- c("State","Cumulative Federal Grants (CWSRF and DWSRF)_Dollars")
  
levels(GDP_RPP$variable); levels(GDP_RPP$Units)
GDP_RPP <- dcast(GDP_RPP, State ~ variable)
colnames(GDP_RPP) <- c("State", "Gross Domestic Product_CurrentDollars","Regional Price Parities_Index")

levels(InterstateCompacts$variable)
InterstateCompacts <- dcast(InterstateCompacts,State~variable)

levels(Labor$variable); levels(Labor$Units)
Labor <- Labor %>%
  filter(variable %in% c("Median Household Income","Unemployment Rate")) %>%
  dcast(State ~ variable)
colnames(Labor) <- c("State","Median Household Income_Infl-adjDollars","Unemployment Rate_percent")

levels(LandUse$variable); levels(LandUse$Units)
LandUse$variable <- paste(LandUse$variable,"percent",sep="_")
LandUse <- dcast(LandUse, State~variable)

levels(LeadingDamOwnership$variable)
LeadingDamOwnership <- dcast(LeadingDamOwnership, State~variable)

levels(NetElectricityGeneration$variable); levels(NetElectricityGeneration$Units)
NetElectricityGeneration$variable <- paste(NetElectricityGeneration$variable,"thousMWh",sep="_")
NetElectricityGeneration <- dcast(NetElectricityGeneration,State~variable)

levels(NutrientTradingPrgrms$variable); levels(NutrientTradingPrgrms$Notes)
NutrientTradingPrgrms <- dcast(NutrientTradingPrgrms, State~variable)

levels(PercentageofHydropowerGeneration$variable)
PercentageofHydropowerGeneration$variable <- paste(PercentageofHydropowerGeneration$variable,
                                                   "percent",sep="_")
PercentageofHydropowerGeneration <- dcast(PercentageofHydropowerGeneration, State~variable)

levels(PopulationSize_Growth$variable);levels(PopulationSize_Growth$Units)
PopulationSize_Growth <- dcast(PopulationSize_Growth,State~variable)

levels(PublicSupply_DW$variable); levels(PublicSupply_DW$Units)
PublicSupply_DW$variable <- paste(PublicSupply_DW$variable,"percent",sep="_")
PublicSupply_DW <- dcast(PublicSupply_DW, State~variable)

levels(PVI$variable);levels(PVI$Units); levels(PVI$Notes)
PVI <- dcast(PVI, State~variable)

levels(RaceEthnicity$variable); levels(RaceEthnicity$Units)
RaceEthnicity$variable <- paste(RaceEthnicity$variable,"percent",sep="_")
RaceEthnicity <- dcast(RaceEthnicity,State~variable)

levels(SewageDisposal1990$variable); levels(SewageDisposal1990$Units)
SewageDisposal1990$variable <- paste(SewageDisposal1990$variable,"percent",sep="_")
SewageDisposal1990 <- dcast(SewageDisposal1990,State~variable)

levels(StateLegislatureParties$variable)
StateLegislatureParties <- dcast(StateLegislatureParties, State~variable)

levels(TotalDamStorage$variable);levels(TotalDamStorage$Units)
TotalDamStorage$variable <- paste(TotalDamStorage$variable,"Ac-ft",sep="_")
TotalDamStorage<- dcast(TotalDamStorage,State~variable)

levels(TotalWaterUse$variable); levels(TotalWaterUse$Units)
TotalWaterUse$variable <-paste(TotalWaterUse$variable,"MGD",sep="_")
TotalWaterUse <- dcast(TotalWaterUse,State~variable) #, value.var = "Total"

levels(UrbanPercent_Growth$variable)
UrbanPercent_Growth <- dcast(UrbanPercent_Growth,State~variable)

levels(WaterPovertyRisk$variable); levels(WaterPovertyRisk$Units)
WaterPovertyRisk$variable <- "WaterPovertyHighAtRisk_%ofTracts"
WaterPovertyRisk <- dcast(WaterPovertyRisk,State~variable)

levels(WaterUse_Sector$variable); levels(WaterUse_Sector$Units)
WaterUse_Sector$Sector <- paste("Percent_Useby",WaterUse_Sector$Sector,sep = "_")
WaterUse_Sector <- dcast(WaterUse_Sector,State~Sector)

levels(WaterUsebySource_Quality$variable)
WaterUsebySource_Quality$Source <- paste("Percent_SalinetoTotal",WaterUsebySource_Quality$Source,
                                         sep = "_")
WaterUsebySource_Quality <- dcast(WaterUsebySource_Quality,State~Source)

levels(WaterUsebySource_SW$variable); levels(WaterUsebySource_SW$Units)
WaterUsebySource_SW$variable <- "Surface water use_percent of total water use"
WaterUsebySource_SW <- dcast(WaterUsebySource_SW,State~variable)

colnames(StateDroughtPlansInfo)
StateDroughtPlansInfo$X <- NULL

#...bind all of the dfs 
AllPredictors <- do.call(cbind, list(Age,AridityIndex_Median_Variability,CleanWaterNeeds,
                                     ClimateInWaterPlans,ConservationScore_ElectricityUse,
                                     Education,FederalGrantsAwarded,GDP_RPP,InterstateCompacts,
                                     Labor,LandUse,LeadingDamOwnership,NetElectricityGeneration,
                                     NutrientTradingPrgrms,PercentageofHydropowerGeneration,
                                     PopulationSize_Growth,PublicSupply_DW,PVI,RaceEthnicity,
                                     SewageDisposal1990,StateLegislatureParties,TotalDamStorage,
                                     TotalWaterUse,UrbanPercent_Growth,WaterPovertyRisk,
                                     WaterUse_Sector,WaterUsebySource_Quality,WaterUsebySource_SW, StateDroughtPlansInfo))

AllPredictors <- AllPredictors[!duplicated(as.list(AllPredictors))] #drop duplicate state label columns

rm(Age,AridityIndex_Median_Variability,CleanWaterNeeds,
                                     ClimateInWaterPlans,ConservationScore_ElectricityUse,
                                     Education,FederalGrantsAwarded,GDP_RPP,InterstateCompacts,
                                     Labor,LandUse,LeadingDamOwnership,NetElectricityGeneration,
                                     NutrientTradingPrgrms,PercentageofHydropowerGeneration,
                                     PopulationSize_Growth,PublicSupply_DW,PVI,RaceEthnicity,
                                     SewageDisposal1990,StateLegislatureParties,TotalDamStorage,
                                     TotalWaterUse,UrbanPercent_Growth,WaterPovertyRisk,
                                     WaterUse_Sector,WaterUsebySource_Quality,WaterUsebySource_SW)

```

###Save data into easily usable files

```{r}
# write.csv(AllPredictors,"../Data/MasterDF/AllPredictors.csv")
# write.csv(Responsedf,"../Data/MasterDF/ResponseVars.csv")
# write.csv(StateInfo,"../Data/MasterDF/StateInfo.csv")
```